### 2016-12-19 results.tex
no message

\begin{theorem}%[{\rm Correctness of linear-PRF}]
\label{thm1}\label{thm:lin-correctness} Fix $k,m,n,r>0$ and~$\elts$,
and let $\setprim_{\mathrm{lin}}= (\Rep, \Qry)$ be the
set-membership data structure over~$\elts$.  For any adversary~$A$
that makes a total of~$q_T$ queries to its $\Test$ oracle and runs
in time~$t$, there exist~$B_1,B_2$ such that
\[
\AdvCorrect{\setprim_{\mathrm{lin}},r}{A} \leq  \AdvPRF{F}{B_1} + \AdvPRF{F}{B_2}  +{\dbinom{q}{r}} \cdot \left( (1-e^{-kn/m})^k + O(1/n) \right)^r\,.
\]
Here, $B_1$ and $B_2$ each ask~$q$ oracle queries and run in
time~$O(t+qm)$.
\end{theorem}
Kirsch and Mitzenmacher~\cite{kirsch2006less} show that the ($r=1$,
zero-query) error-probability for~$\setprim_\mathrm{lin}$ is always
within $O(1/n)$ of $(1-e^{-kn/m})^k$, although the asymptotic rate
of convergence is faster than~$1/n$; the multiplicative term
$\binom{q}{r}$ arises due to switching from the zero-query setting
to the adaptive setting. A proof of the above will appear in the
full version, but we remark that it closely follows the proof of
Theorem~\ref{thm:bf-correctness}.
%\input{lin-correctness}

\heading{Privacy. } Simulation-based privacy of the linear PRF-hash
construction follows from~Theorem~\ref{thm:priv-ss}.

@@

%%%%%%%%%%%%%%%%% BIGRAM LINEAR-PRF (AKA "NIEDERMEYER")  %%%%%%%%%%%%%%%%%%%%
\subsection{Bigram Linear-PRF Construction}
%\heading{Bigram Linear-PRF construction (``Niedermeyer''). }
%\heading{Correctness.}
%\todo{Correctness theorem for this scheme.  Should follow pretty closely the one for ``plain'' linear-PRF.}\tsnote{Actually not sure this is that interesting, but for parallel structure one would want to give it.}
Consider the set-membership data structure given in
Figure~\ref{fig:lin-and-bi-lin} (right side), which we call the
\emph{bigram linear-PRF} construction. \jnote{Really a bad
name\ldots} This construction is of interest as it has been
suggested as a way to address the privacy-preserving record-linkage
problem, in particular for linking common records in medical
databases~\cite{niedermeyer2014cryptanalysis}. %,schnell2011novel}.
\jnote{We should really give more discussion, explaining the data
structure in words, and also discussing the goal of
``privacy-preserving record linkage'' and describing the protocol
for doing so (at a high level).}

\heading{Privacy. } We only address the privacy of this
construction, as the correctness follows immediately from the
correctness of the linear PRF-hash construction. \jnote{Well, with
some loss of parameters.} The privacy, however, does not.
%JK
Here we
exhibit an efficient SS-privacy attack even when simulator is
provided nearly the entire set. Consider the following
``all-but-last'' leakage function. On input a
set~$S=\{x_1,x_2,\ldots,x_n\}$, it assigns $x'_n \gets
x_n[1:|x_n|-1]$ and returns
$\mathsf{leak_{abl}}(S)=\{x_1,x_2,\ldots,x'_n\}$.

\begin{figure}[tp]
\centering
\fpage{.65}
{
\adversaryv{$A$}\\[1ex]
On input $\emptystring$:\\
\nudge $x_1 \gets \mathrm{cab}$\\
\nudge $x_2 \gets \mathrm{bab}$\\
\nudge $z \getsr \bits$\\
\nudge if $z=1$ then $x_3 \gets \mathrm{caba}$\\
\nudge else $x_3 \gets \mathrm{cabc}$\\
\nudge Return $S=\{x_1,x_2,x_3\}$\\
On input $\pub \neq \emptystring$:\\
\nudge $M_1,M_2,M_3 \gets \pub$\\
\nudge $M_{1,2} \gets M_1 \vee M_2$\\
\nudge if $z=1 \wedge \left(3k < \mathsf{hamming}(M_{1,2} \wedge M_3) = 4k\right)$ then \\
\nudge\nudge Return 1 \\
\nudge if $z=0 \wedge \left(\mathsf{hamming}(M_{1,2} \wedge M_3) \leq 3k\right)$ then \\
\nudge\nudge Return 1 \\
\nudge Return 0
}
\caption{Adversary for Theorem~\ref{thm:bi-lin-ss}. The function $\mathsf{hamming}(M)$ returns the hamming weight of~$M$.}
\label{fig:adv-bi-lin-ss}
\end{figure}

\begin{theorem}[{\rm SS-privacy of bigram linear-PRF}]\label{thm:bi-lin-ss}
Let $\setprim_{\bilin}= (\Rep, \Qry)$ be the set-membership data
structure described in Figure~\ref{fig:lin-and-bi-lin}, but with
$F_{K}$ replaced by a random function~$\rho$. Let
$\leak_{\mathsf{abl}}$ be the leakage function just described.
Let~$\elts=\{\mathrm{a,b,c}\}^*$. Let $n=3$,  $m=3\alpha $ for
integer $\alpha >> 1$, and let, $k = \lceil \ln(2) \alpha
\rceil$.\footnote{This value of~$k$ is from the classical analysis for
minimizing the FP rate for a Bloom filter, and is set this way for
compatibility with classical results.  It isn't used in any
essential way in the proof sketch.}  Let~$A$ be the adversary in
Figure~\ref{fig:adv-bi-lin-ss}.  Then for any simulator $\Sim$, the
advantage
$\AdvPrivSS{\setprim_{\bilin},\leak_{\mathsf{abl}}}{A,\Sim}$ can be
made arbitrarily close to $1/2$ by choice of~$\alpha$.
\end{theorem}
\begin{proof}%\tsnote{still a bit sketchy but believable}
On input the set~$S$ produced by~$A$, the $\Rep$ algorithm turns $x_1=$'cab', $x_2=$'bab' and $x_3$ into their respective bigram sets: $B_{x_1}=\{\sqcup\mathrm{c},\mathrm{ca},\mathrm{ab},\mathrm{b}\sqcup\}$, $B_{x_2}=\{\sqcup\mathrm{b},\mathrm{ba},\mathrm{ab}, \mathrm{b}\sqcup \}$; the set $B_{x_3}$ will be either $\{\sqcup\mathrm{c},\mathrm{ca},\mathrm{ab},\mathrm{ba}, \mathrm{a}\sqcup\}$, or $\{\sqcup\mathrm{c},\mathrm{ca},\mathrm{ab},\mathrm{bc}, \mathrm{c}\sqcup\}$, each with probability 1/2.  It then produces $\alpha$-bit arrays $M_1, M_2, M_3$ by hashing the corresponding bigram sets using the hash functions $h_j(s)=\rho_1(s)+ j\rho_2(s) \bmod m$.  Note that for sufficiently large~$m$, each of $M_1,M_2$ will contain exactly $4k$ positions set to 1; this is with overwhelming probability, up to collisions in the hash functions for different bigrams within a given bigram set.  Likewise, $M_{1,2}=M_1 \vee M_2$ will have exactly $6k$ positions set to 1, because $B_{x_1}$ and $B_{x_2}$ share $\mathrm{ab}$ and $\mathrm{a}\sqcup$.

In the case that $B_{x_3}=\{\sqcup\mathrm{c},\mathrm{ca},\mathrm{ab},\mathrm{ba}, \mathrm{a}\sqcup\}$, only one of its elements ($\mathrm{a}\sqcup$) is not in either of $B_{x_1},B_{x_2}$.  Hence, the hamming weight of $M_{1,2}\wedge M_3$ will be $4k$.  On the other hand, when $B_{x_3} = \{\sqcup\mathrm{c},\mathrm{ca},\mathrm{ab},\mathrm{bc}, \mathrm{c}\sqcup\}$, two of its elements ($\mathrm{bc}, \mathrm{c}\sqcup$) are not covered by $B_{x_1},B_{x_2}$.  In this case, the hamming weight of $M_{1,2} \wedge M_3$ will be $3k$.  Thus $\ExpPrivSSreal{\setprim_{\bilin}}{A}=1$ with overwhelming probability.

On the other hand, the simulator~$\Sim$ is given 'cab','bab','cab' as
input.  It can perfectly simulate $M_1,M_2$ by lazily sampling random
function~$\rho$ itself.  However, it has no information about the
bit~$z$, so at best it can cause~$A$ to output 1 with
probability~$1/2$.  That is $\ExpPrivSSsim{\setprim_{\bilin}}{A,\Sim} \leq 1/2$. \hfill\qed
\end{proof}

>>>>

\begin{theorem}%[{\rm Correctness of linear-PRF}]
\label{thm1}\label{thm:lin-correctness} Fix $k,m,n,r>0$ and~$\elts$,
and let $\setprim_{\mathrm{lin}}= (\Rep, \Qry)$ be the
set-membership data structure over~$\elts$.  For any adversary~$A$
that makes a total of~$q_T$ queries to its $\Test$ oracle and runs
in time~$t$, there exist~$B_1,B_2$ such that
\[
\AdvCorrect{\setprim_{\mathrm{lin}},r}{A} \leq  \AdvPRF{F}{B_1} + \AdvPRF{F}{B_2}  +{\dbinom{q}{r}} \cdot \left( (1-e^{-kn/m})^k + O(1/n) \right)^r\,.
\]
Here, $B_1$ and $B_2$ each ask~$q$ oracle queries and run in
time~$O(t+qm)$.
\end{theorem}
Kirsch and Mitzenmacher~\cite{kirsch2006less} show that the ($r=1$,
zero-query) error-probability for~$\setprim_\mathrm{lin}$ is always
within $O(1/n)$ of $(1-e^{-kn/m})^k$, although the asymptotic rate
of convergence is faster than~$1/n$; the multiplicative term
$\binom{q}{r}$ arises due to switching from the zero-query setting
to the adaptive setting. A proof of the above closely follows the
proof of Theorem~\ref{thm:bf-correctness}, and will appear in the
full version.
%\input{lin-correctness}

\heading{Privacy. } Simulation-based privacy of the linear PRF-hash
construction follows from~Theorem~\ref{thm:priv-ss}:

@@

%%%%%%%%%%%%%%%%% BIGRAM LINEAR-PRF (AKA "NIEDERMEYER")  %%%%%%%%%%%%%%%%%%%%
\subsection{Bigram Linear-PRF Construction}
%\heading{Bigram Linear-PRF construction (``Niedermeyer''). }
%\heading{Correctness.}
%\todo{Correctness theorem for this scheme.  Should follow pretty closely the one for ``plain'' linear-PRF.}\tsnote{Actually not sure this is that interesting, but for parallel structure one would want to give it.}

The set-membership data structure given on the right-hand side of
Figure~\ref{fig:lin-and-bi-lin}, which we call the \emph{bigram
linear-PRF} construction \jnote{Really a bad name\ldots}, was
suggested as a way to allow for privacy-preserving
linkage of medical records~\cite{niedermeyer2014cryptanalysis}. %,schnell2011novel}.
\jnote{We should really give more discussion, explaining what the
data structure does in words, and also discussing the goal of
``privacy-preserving record linkage'' and describing the protocol
for doing so (at a high level).} We only address the privacy of this
construction, as the correctness follows immediately from the
correctness of the linear PRF-hash construction. \jnote{Actually, it
seems worth giving a proof of correctness just to get the parameters
right.}

\heading{Privacy. }  Unfortunately, and in contrast to the linear
PRF-hash construction on which it is based, the scheme does not
achieve our simulation-based notion of privacy. In fact, we show
that it fails severely: roughly, even a simulator given everything
but the last character of one of the strings in~$S$ cannot simulate
the public output.

%\begin{figure}[tp]
%\centering
%\fpage{.65}
%{
%\adversaryv{$A$}\\[1ex]
%On input $\emptystring$:\\
%\nudge $x_1 \gets \mathrm{cab}$\\
%\nudge $x_2 \gets \mathrm{bab}$\\
%\nudge $z \getsr \bits$\\
%\nudge if $z=1$ then $x_3 \gets \mathrm{caba}$\\
%\nudge else $x_3 \gets \mathrm{cabc}$\\
%\nudge Return $S=\{x_1,x_2,x_3\}$\\
%On input $\pub \neq \emptystring$:\\
%\nudge $M_1,M_2,M_3 \gets \pub$\\
%\nudge $M_{1,2} \gets M_1 \vee M_2$\\
%\nudge if $z=1 \wedge \left(3k < \mathsf{hamming}(M_{1,2} \wedge M_3) = 4k\right)$ then \\
%\nudge\nudge Return 1 \\
%\nudge if $z=0 \wedge \left(\mathsf{hamming}(M_{1,2} \wedge M_3) \leq 3k\right)$ then \\
%\nudge\nudge Return 1 \\
%\nudge Return 0
%}
%\caption{Adversary for Theorem~\ref{thm:bi-lin-ss}. The function $\mathsf{hamming}(M)$ returns the hamming weight of~$M$.}
%\label{fig:adv-bi-lin-ss}
%\end{figure}

\jnote{Give intuition for why the attack works: the same ``hash
functions'' are used for each ``Bloom filter,'' and this leaks
information about bigrams shared among multiple strings.} \jnote{I
simplified the attack.}

Formally, consider sets of strings over $\bool$, and fix the leakage
function $\leak_{\mathsf{abl}}$ defined by
\[\leak_{\mathsf{abl}}(\{x_1, \ldots, x_n\}) = \{x_1, \ldots, x_{n-1},
x'_n\},\] where $x_n$ is the lexicographically last string in the
given set, and $x'_n$ is equal to $x_n$ with the last bit removed.
Define an attacker $A$ as follows: it chooses a uniform bit~$b$; if
$b=0$ then it outputs the set $S=\{1,111\}$, and if $b=1$ then it
outputs the set $S=\{1, 101\}$. Next, given public information $\pub
= \langle M_1, M_2 \rangle$ (where $M_2$ is the array with more bits
set to~1), it counts the number of positions set to~1 in $M_2$ that
are not set to~1 in~$M_1$. If this number is strictly greater
than~$k$, then $A$ outputs~1; otherwise, it outputs~0.

In analyzing the behavior of $A$, note first that $\Rep$ begins by
mapping each of the strings in $S$ to their respective bigram sets,
namely, $\{\sqcup 1, 1\sqcup\}$ for the first string, and either
$\{\sqcup 1, 11, 1\sqcup\}$ or $\{\sqcup 1, 10, 01, 1\sqcup\}$ for
the second string. It then produces the $m$-bit arrays $M_1, M_2$.
Because the bigram set for the second string is, in either case, a
strict superset of the bigram set for the first string, $M_2$ (which
is the array with more bits set to~1) always corresponds to the
second string.

Now, when $b=0$ the bigram set for the second string contains
exactly one more element than the bigram set for the first string,
and so the number of positions set to~1 in $M_2$ that are \emph{not}
set to~1 in $M_1$ is at most~$k$. On the other hand, when $b=1$ the
bigram set for the second string contains exactly two more elements
than the bigram set for the first string, and so with high
probability the number of positions set to~1 in $M_2$ that are not
set to~1 in $M_1$ exceeds~$k$. (Specifically, in the worst case
there are $2k$ positions---call them~$I$---set to~1 in $M_1$, and
then the number in question exceeds~$k$ as long as a strict majority
of the $2k$ positions set to~1 by the extra bigrams $10, 01$ lie
outside of~$I$. Pairwise independence of those $2k$ positions can be
used to bound this probability. If $k\geq 4$ and $2k/m \leq 1/4$,
this probability is at least %JK

 implies that this happens except with
probability at most $(4 \cdot (\frac{1}{2}- \leq 1/2$.)

%\begin{theorem}%[{\rm SS-privacy of bigram linear-PRF}]
%\label{thm:bi-lin-ss}
%Let~$\elts=\{\mathrm{a,b,c}\}^*$. Let $n=3$,  $m=3\alpha $ for
%integer $\alpha \gg 1$, and let, $k = \lceil \ln(2) \alpha
%\rceil$.\footnote{This value of~$k$ is from the classical analysis
%for minimizing the FP rate for a Bloom filter, and is set this way
%for compatibility with classical results.  It isn't used in any
%essential way in the proof sketch.}  Let~$A$ be the adversary in
%Figure~\ref{fig:adv-bi-lin-ss}.  Then for any simulator $\Sim$, the
%advantage
%$\AdvPrivSS{\setprim_{\bilin},\leak_{\mathsf{abl}}}{A,\Sim}$ can be
%made arbitrarily close to $1/2$ by choice of~$\alpha$.
%\end{theorem}
\begin{proof}%\tsnote{still a bit sketchy but believable}
On the other hand, the simulator~$\Sim$ is given 'cab','bab','cab'
as input.  It can perfectly simulate $M_1,M_2$ by lazily sampling
random function~$\rho$ itself.  However, it has no information about
the bit~$z$, so at best it can cause~$A$ to output 1 with
probability~$1/2$.  That is
$\ExpPrivSSsim{\setprim_{\bilin}}{A,\Sim} \leq 1/2$. \hfill\qed
\end{proof}
